### 2023-07-16 tex/part.tex
[ writing ] address all remaining todos

As we have seen, the way additive and multiplicative rules are
realised algebraically is related to models of separation logic.
Models of separation logic typically use \emph{partial} commutative monoids to
model a heap, so it is tempting to generalise the commutative monoid of
addition in our semirings to a \emph{partial} commutative monoid.
However, we find that the most natural notion of \emph{partial semiring} is
degenerate, in that all partial semirings are actually (total) semirings.

Recall that a commutative monoid (or commutative monoid object) can be
defined in any symmetric monoidal category.
A partial commutative monoid is exactly a commutative monoid object in the
category of sets and partial functions with the usual monoidal product
\todo{Define this monoidal product somewhere.}.
However, semirings need a Cartesian category in order to state the interaction
equations between addition and multiplication.
While the category of sets and partial functions is not Cartesian, the
standard way to manufacture a Cartesian category out of a symmetric monoidal
category $\mathcal C$ is to take the category of cocommutative comonoids
$\mathrm{CComon}(\mathcal C)$.
Intuitively, the cocommutative comonoid structure equips the underlying
object $M$ with a \emph{delete} map $\eta : M \to I$ and a \emph{duplicate}
map $\delta : M \to M \otimes M$ which are coherent with respect to each other.
All morphisms in $\mathrm{CComon}(\mathcal C)$ must respect $\eta$ and
$\delta$; in particular, both addition and multiplication must separately
form bimonoids in $\mathcal C$ together with the cocommutative comonoid.

@@

    \draw (0) -- (*);
    \draw (x) to[out=270,in=45] (*);
    \draw (*) -- (res);
  \end{tikzpicture}
  =\quad
  \begin{tikzpicture}[baseline]
    \path
    (0,0) node(0) {0}
    (0,2) node(x) {}
    (0,-1) node(res) {}
    (0,1) node[circle,draw](del) {}
    ;

@@

      \draw (x) to[out=270,in=135] (+);
      \draw (y) to[out=270,in=45] (+);
      \draw (+) to[out=270,in=135] (*);
      \draw (z) to[out=270,in=45] (*);
      \draw (*) -- (res);
    \end{tikzpicture}
    =
    \begin{tikzpicture}[baseline]
      \path
      (-2,3) node(x) {}
      (-1,3) node(y) {}
      (0,3) node(z) {}
      (0,2) node[circle,draw](dup) {}
      (-1,1) node(x*) {*}
      (0,1) node(y*) {*}
      (-0.5,0) node(+) {+}
      (-0.5,-1) node(res) {}
      ;

      \draw (z) -- (dup);
      \draw (x) to[out=270,in=135] (x*);
      \draw (y) to[out=270,in=135] (y*);
      \draw (dup) to[out=180,in=45] (x*);
      \draw (dup) -- (y*);
      \draw (x*) to[out=270,in=135] (+);
      \draw (y*) to[out=270,in=45] (+);
      \draw (+) -- (res);
    \end{tikzpicture}
    &\phantom{mmmm}&
    \begin{tikzpicture}[baseline]
      \path
      (1,2) node(x) {}
      (0,2) node(y) {}
      (0.5,1) node(+) {+}
      (-1,2) node(z) {}
      (0,0) node(*) {*}
      (0,-1) node(res) {}
      ;

      \draw (x) to[out=270,in=45] (+);
      \draw (y) to[out=270,in=135] (+);
      \draw (+) to[out=270,in=45] (*);
      \draw (z) to[out=270,in=135] (*);
      \draw (*) -- (res);
    \end{tikzpicture}
    =
    \begin{tikzpicture}[baseline]
      \path
      (2,3) node(x) {}
      (1,3) node(y) {}
      (0,3) node(z) {}
      (0,2) node[circle,draw](dup) {}
      (1,1) node(x*) {*}
      (0,1) node(y*) {*}
      (0.5,0) node(+) {+}
      (0.5,-1) node(res) {}
      ;

      \draw (z) -- (dup);
      \draw (x) to[out=270,in=45] (x*);
      \draw (y) to[out=270,in=45] (y*);
      \draw (dup) to[out=0,in=135] (x*);
      \draw (dup) -- (y*);
      \draw (x*) to[out=270,in=45] (+);
      \draw (y*) to[out=270,in=135] (+);
      \draw (+) -- (res);
    \end{tikzpicture}
  \end{matrix}
\end{displaymath}

\begin{lemma}
  For each object $X$ in $(\mathrm{Set}_{\mathrm{part}}, {\otimes})$, there is
  a cocommutative comonoid over $X$.
\end{lemma}
\begin{proof}
  Let $\eta(x) \coloneq ()$ and $\delta(x) \coloneq (x, x)$, with both
  being defined for all $x$.
  Checking that these satisfy the laws is routine.
  Alternatively, we can see that both $\eta$ and $\delta$, being total, are
  morphisms in $\mathrm{Set}$, where it is well known that they form a
  cocommutative comonoid.
  The equations in $\mathrm{Set}$ carry over to $\mathrm{Set}_{\mathrm{part}}$.
\end{proof}

\begin{lemma}
  For each object $X$ in $(\mathrm{Set}_{\mathrm{part}}, {\otimes})$, any
  comonoid over $X$ is the one described in the previous lemma.
\end{lemma}
\begin{proof}
  The left unit law says that, for all $x$ and $x'$, we have
  $\exists y.~\delta(x) = (y, x') \land \eta(y) = ()$ if and only if $x = x'$.
  Letting $x'$ be $x$ and reading from right to left, we get that there is
  some $y$ such that $\delta(x) = (y, x)$ and $\eta(y) = ()$.
  Symmetrically, from the right unit law, there is some $z$ such that
  $\delta(x) = (x, z)$ and $\eta(z) = ()$.
  But because $\delta$, as a partial function, is deterministic, we have
  $(y, x) = (x, z)$, giving us that $y = z = x$, and $\delta(x) = (x, x)$.
  Moreover, $\eta(x) = \eta(y) = ()$.
\end{proof}

>>>>

As we have seen, the way additive and multiplicative rules are
realised algebraically is related to models of separation logic.
Models of separation logic typically use \emph{partial} commutative monoids to
model a heap, so it is tempting to generalise the commutative monoid of
addition in our semirings to a \emph{partial} commutative monoid.
However, we find that the most natural notion of \emph{partial semiring} is
degenerate, in the sense that all partial semirings are actually (total)
semirings.

Recall that a commutative monoid (or commutative monoid object) can be
defined in any symmetric monoidal category.
A partial commutative monoid is exactly a commutative monoid object in the
category of sets and partial functions with the usual monoidal product given
by pairing of objects and morphisms (like the Cartesian product in $\Set$).
However, semirings need a Cartesian category in order to state the interaction
equations between addition and multiplication.
While the category of sets and partial functions is not Cartesian, the
standard way to manufacture a Cartesian category out of a symmetric monoidal
category $\mathcal C$ is to take the category of cocommutative comonoids
$\mathrm{CComon}(\mathcal C)$.
Intuitively, the cocommutative comonoid structure equips the underlying
object $M$ with a \emph{delete} map $\eta : M \to I$ and a \emph{duplicate}
map $\delta : M \to M \otimes M$ which are coherent with respect to each other.
All morphisms in $\mathrm{CComon}(\mathcal C)$ must respect $\eta$ and
$\delta$; in particular, both addition and multiplication must separately
form bimonoids in $\mathcal C$ together with the cocommutative comonoid.

The distributivity laws of semirings are stated below.
I include these to show that the cocommutative comonoids of a monoidal category
give enough structure to state these laws.
The other laws --- that all morphisms respect $\eta$ and $\delta$, that addition
forms a commutative monoid, and that multiplication forms a monoid --- are
standard in symmetric monoidal category theory.

@@

    \draw (0) -- (*);
    \draw (x) to[out=270,in=45] (*);
    \draw (*) -- (res);
  \end{tikzpicture}
  =\quad
  \begin{tikzpicture}[baseline]
    \path
    (0,0) node(0) {0}
    (0,2) node(x) {}
    (0,-1) node(res) {}
    (0,1) node(del) {$\eta$}
    ;

@@

      \draw (x) to[out=270,in=135] (+);
      \draw (y) to[out=270,in=45] (+);
      \draw (+) to[out=270,in=135] (*);
      \draw (z) to[out=270,in=45] (*);
      \draw (*) -- (res);
    \end{tikzpicture}
    =
    \begin{tikzpicture}[baseline]
      \path
      (-2,3) node(x) {}
      (-1,3) node(y) {}
      (0,3) node(z) {}
      (0,2) node(dup) {$\delta$}
      (-1,1) node(x*) {*}
      (0,1) node(y*) {*}
      (-0.5,0) node(+) {+}
      (-0.5,-1) node(res) {}
      ;

      \draw (z) -- (dup);
      \draw (x) to[out=270,in=135] (x*);
      \draw (y) to[out=270,in=135] (y*);
      \draw (dup) to[out=-150,in=45] (x*);
      \draw (dup) -- (y*);
      \draw (x*) to[out=270,in=135] (+);
      \draw (y*) to[out=270,in=45] (+);
      \draw (+) -- (res);
    \end{tikzpicture}
    &\phantom{mmmm}&
    \begin{tikzpicture}[baseline]
      \path
      (1,2) node(x) {}
      (0,2) node(y) {}
      (0.5,1) node(+) {+}
      (-1,2) node(z) {}
      (0,0) node(*) {*}
      (0,-1) node(res) {}
      ;

      \draw (x) to[out=270,in=45] (+);
      \draw (y) to[out=270,in=135] (+);
      \draw (+) to[out=270,in=45] (*);
      \draw (z) to[out=270,in=135] (*);
      \draw (*) -- (res);
    \end{tikzpicture}
    =
    \begin{tikzpicture}[baseline]
      \path
      (2,3) node(x) {}
      (1,3) node(y) {}
      (0,3) node(z) {}
      (0,2) node(dup) {$\delta$}
      (1,1) node(x*) {*}
      (0,1) node(y*) {*}
      (0.5,0) node(+) {+}
      (0.5,-1) node(res) {}
      ;

      \draw (z) -- (dup);
      \draw (x) to[out=270,in=45] (x*);
      \draw (y) to[out=270,in=45] (y*);
      \draw (dup) to[out=-30,in=135] (x*);
      \draw (dup) -- (y*);
      \draw (x*) to[out=270,in=45] (+);
      \draw (y*) to[out=270,in=135] (+);
      \draw (+) -- (res);
    \end{tikzpicture}
  \end{matrix}
\end{displaymath}

It is well known that all commutative comonoids in $(\Set, \times)$, and indeed
any Cartesian monoidal category, are trivial, in the sense that every object of
$\Set$ gives rise to exactly one commutative comonoid.
We find in the following two lemmas that this property also holds of
$\plr{\Set_{\mathrm{part}}, \otimes}$.

\begin{lemma}\label{thm:ccomon-exists}
  For each object $X$ in $\plr{\Set_{\mathrm{part}}, {\otimes}}$, there is
  a cocommutative comonoid over $X$.
\end{lemma}
\begin{proof}
  Let $\eta(x) \coloneq ()$ and $\delta(x) \coloneq (x, x)$, with both
  being defined for all $x$.
  Checking that these satisfy the cocomutative comonoid laws is routine.
  Alternatively, we can see that both $\eta$ and $\delta$, being total, are
  morphisms in $\mathrm{Set}$, where it is well known that they form a
  cocommutative comonoid.
  The equations in $\mathrm{Set}$ carry over to $\mathrm{Set}_{\mathrm{part}}$.
\end{proof}

\begin{lemma}\label{thm:ccomon-unique}
  For each object $X$ in $\plr{\Set_{\mathrm{part}}, {\otimes}}$, any
  comonoid over $X$ is the one described in \cref{thm:ccomon-exists}.
\end{lemma}
\begin{proof}
  The left unit law says that, for all $x$ and $x'$, we have
  $\exists y.~\delta(x) = (y, x') \land \eta(y) = ()$ if and only if $x = x'$.
  Letting $x'$ be $x$ and reading from right to left, we get that there is
  some $y$ such that $\delta(x) = (y, x)$ and $\eta(y) = ()$.
  Symmetrically, from the right unit law, we get some $z$ such that
  $\delta(x) = (x, z)$ and $\eta(z) = ()$.
  But because $\delta$, being a partial function, is deterministic, we have
  $(y, x) = (x, z)$, giving us that $y = z = x$, and $\delta(x) = (x, x)$.
  Moreover, because the chosen $y$ is equal to $x$, we have for all $x$ that
  $\eta(x) = ()$.
\end{proof}

That a morphism $f$ respects the $\eta$ given in \cref{thm:ccomon-exists} is
equivalent to saying that $f$ is total.
Therefore, all possible semiring operators in
$\mathrm{CComon}\plr{\Set_{\mathrm{part}}, \otimes}$ are total, meaning that
there is a corresponding semiring in $\plr{\Set, \times}$.

The above reasoning shows that semirings in the category of sets and partial
functions are not worth studying.
If we want partiality, there appear to be two options.
The first option is to give up on multiplication.
We could imagine replacing the binary multiplication operator by a set of
unary modalities satisfying fewer laws.
In particular, I make little use of addition on the left of a multiplication,
and multiplying by $\gr0$ on the left (as done by $\oc\gr0$) is unwanted in some
cases (such as when encoding DILL and PD, as in \cref{sec:translation}).
With unary modalities, we could expect all of the required laws to be
expressible in a symmetric monoidal category.
The second option is to use a different notion of partiality.
The notion of partiality given by the category of sets and partial functions is
``strict'', in that composing with an everywhere-undefined function yields an
everywhere-undefined function.
With a non-strict notion of partial function, we may be able to have interesting
partial semirings.

### 2023-07-16 tex/semantics.tex
[ writing ] various up to WRel example
[ writing ] address all remaining todos

This chapter is structured as follows.
I start by giving a quick introduction to linear relations --- a generalisation
of linear maps --- with reference to their use in mechanised algebraic
reasoning, in \cref{sec:lin-rel}.
Using linear relations, I give a functorial \emph{map} operation to a single
layer of syntax in \cref{sec:functorial}.
I then adapt the \AgdaFunction{Kripke} function space to the usage-aware
setting in \cref{sec:kripke}.
Then I apply the \AgdaFunction{Kripke} function space, along with much of the
machinery I have introduced in previous chapters, to give the
\AgdaFunction{semantics} function in \cref{sec:traversal}.
Finally, I give some example uses of \AgdaFunction{semantics} in
\cref{sec:example-semantics}.
\todo{What are the examples?}

@@

The linearity of relation \AgdaBound{$\Psi$} is given by fields
\AgdaField{rel-0$_m$}, \AgdaField{rel-+$_m$}, and \AgdaField{rel-*$_m$} (with
the subscript-m being a mnemonic for \emph{module}, as opposed to scalar).
\todo{Refer to preliminaries}

@@

At this point I introduce a minor generalisation to
\AgdaFunction{OpenFam} and \AgdaFunction{ExtOpenFam} (as defined in
\cref{sec:terms}):
\AgdaBound{I}\AgdaSpace{}\AgdaFunction{---OpenFam} and
\AgdaBound{I}\AgdaSpace{}\AgdaFunction{---ExtOpenFam}.  We obtain the
definition of \AgdaBound{I}\AgdaSpace{}\AgdaFunction{---OpenFam} by
replacing the textual occurrence of \AgdaBound{Ty} by the parameter
\AgdaBound{I}, though there is still reference to the ambient \AgdaBound{Ty}
via \AgdaRecord{Ctx}.
The main value I am interested in \AgdaBound{I} taking, other than
\AgdaBound{Ty}, is \AgdaRecord{Ctx} --- for example, the type family of
$\V$-environments, for a given $\V$, is a
\AgdaRecord{Ctx}\AgdaSpace{}\AgdaFunction{---OpenFam}%
\AgdaSpace{}\AgdaSymbol{\_}.
However, I only use this generalisation much later.
\todo{When?}

@@

If $\Delta$ is of the form $\gr{s_1}B_1, \ldots, \gr{s_n}B_n$, then
\ExecuteMetaData[\Snippetstex]{KripkeVCDGA}\ is equivalent to
\ExecuteMetaData[\Snippetstex]{KripkeExpanded}\ by Currying.  That is
to say, the Kripke function is expecting a value for each newly bound
variable, at the multiplicity of its annotation, together with the
resources supporting each of those values. We use the ``magic wand''
function space here to enforce the invariant that the freshly bound
variables have usage annotations that are added to the existing
variables, not shared with them. The use of the
\AgdaFunction{$\Box^r$} modality ensures that we can still use it in
the presence of additional variables introduced by weakening.
\todo{Is wand defined?}

@@

\section{Reifying the Kripke function space}

>>>>

This chapter is structured as follows.
I start by giving a quick introduction to linear relations --- a generalisation
of linear maps --- with reference to their use in mechanised algebraic
reasoning, in \cref{sec:lin-rel}.
Using linear relations, I give a functorial \emph{map} operation to a single
layer of syntax in \cref{sec:functorial}.
I then adapt the \AgdaFunction{Kripke} function space to the usage-aware
setting in \cref{sec:kripke}.
Then I apply the \AgdaFunction{Kripke} function space, along with much of the
machinery I have introduced in previous chapters, to give the
\AgdaFunction{semantics} function in \cref{sec:traversal}.
Finally, I give the syntax-generic simultaneous renaming and substitution
operations in \cref{sec:kit-to-sem}.

@@

The linearity of relation \AgdaBound{$\Psi$} is given by fields
\AgdaField{rel-0$_m$}, \AgdaField{rel-+$_m$}, and \AgdaField{rel-*$_m$} (with
the subscript-m being a mnemonic for \emph{module}, as opposed to scalar).

@@

At this point I introduce a minor generalisation to
\AgdaFunction{OpenFam} and \AgdaFunction{ExtOpenFam} (as defined in
\cref{sec:terms-of-system}):
\AgdaBound{I}\AgdaSpace{}\AgdaFunction{---OpenFam} and
\AgdaBound{I}\AgdaSpace{}\AgdaFunction{---ExtOpenFam}.  We obtain the
definition of \AgdaBound{I}\AgdaSpace{}\AgdaFunction{---OpenFam} by
replacing the textual occurrence of \AgdaBound{Ty} by the parameter
\AgdaBound{I}, though there is still reference to the ambient \AgdaBound{Ty}
via \AgdaRecord{Ctx}.
The main value I am interested in \AgdaBound{I} taking, other than
\AgdaBound{Ty}, is \AgdaRecord{Ctx} --- for example, the type family of
$\V$-environments, for a given $\V$, is a
\AgdaRecord{Ctx}\AgdaSpace{}\AgdaFunction{---OpenFam}%
\AgdaSpace{}\AgdaSymbol{\_}.
I use this generality in the type of \AgdaFunction{extend} in the next section.

@@

If $\Delta$ is of the form $\gr{s_1}B_1, \ldots, \gr{s_n}B_n$, then
\ExecuteMetaData[\Snippetstex]{KripkeVCDGA}\ is equivalent to
\ExecuteMetaData[\Snippetstex]{KripkeExpanded}\ by Currying.  That is
to say, the Kripke function is expecting a value for each newly bound
variable, at the multiplicity of its annotation, together with the
resources supporting each of those values. We use the ``magic wand''
function space here to enforce the invariant that the freshly bound
variables have usage annotations that are added to the existing
variables, not shared with them. The use of the
\AgdaFunction{$\Box^r$} modality ensures that we can still use it in
the presence of additional variables introduced by weakening.

@@

\section{Reifying the Kripke function space}\label{sec:reify}

### 2023-07-16 tex/semirings.tex
[ writing ] address all remaining todos

As explained by \citet{McBride16}, the first insight is to, instead of
removing variables from the context of certain subterms, add an annotation to
free variables saying whether or not they are to be used.
I use an annotation $\gr0$ on variables that are not to be used, and an
annotation $\gr1$ on variables that are to be used.
This convention lets us transcribe the usual $\otimes$-introduction rule
(below left) as a rule with usage annotations (below right).
In the notation on the right, I let $\Gamma = \grP\gamma$ and
$\Delta = \grQ\delta$, where $\Gamma$ is a whole context comprising a
\emph{usage context} $\grP$ and a \emph{typing context} $\gamma$.
A usage context is a list of usage annotations, so
$\grP = \gr{r_1}, \ldots, \gr{r_m}$ and a typing context is a list of types, so
$\gamma = A_1, \ldots, A_m$.
When combined, the usage context and the typing context will be of the same
length.
Explicit contexts will usually be written with usage annotations and types
interspersed, as $\gr{r_1}A_1, \ldots, \gr{r_m}A_m$.
I use $\gr r\gamma$ to abbreviate $\gr rx_1, \ldots, \gr rx_m$.

@@

The rest of this chapter proceeds as follows.
In \cref{sec:lr}, I define the usage-annotated calculus \name{}, which has
appeared in my previous work~\citep{WA21}, and can be seen as a
simply typed version of Atkey's dependently typed calculus QTT~\citep{Atkey18}.
The idea of \name{} is to augment the simply typed $\lambda$-calculus with
annotations on free variables, and give enough types to manipulate these
annotations.
Given this new calculus $\name$, the first goal is to apply the techniques of
\cref{sec:simple} to it, yielding a simultaneous substitution operation.
To do this, I use \cref{sec:lnd} to introduce notation that allows us to restate
the typing rules of \name{} to not mention contexts explicitly, as was the style
in \cref{sec:simple}.
This new notation --- the \emph{bunched connectives} --- is versatile at
defining simply typed usage-aware syntaxes, and I give further non-$\name$
examples in \cref{sec:rec}.
Finally, I justify connections to linear logic and modal logic in
\cref{sec:translation}, where I translate $\name$ terms to and from
DILL~\citep{Barber1996} and the modal calculus of \citet{judgmental}.
%The application of the techniques of \cref{sec:simple} is left to
%\cref{sec:ren-sub-lr}.

@@

An important point to note is that $\name$ cannot be instantiated to become a
bunched logic, in the sense of \citet{oHP99}.
This is despite the use of the bunched connectives in the \emph{definition} of
$\name$, and despite the formally tree-shaped contexts.
However, the variable rule of $\name$ essentially treats the variables in the
context independently, except for the individual checking of usage annotations
on each variable.
Therefore, we cannot talk about more interesting spatial relationships between
variables, as required by bunched logic.
This property also precludes us from capturing calculi without exchange
(non-commutative logics), such as Lambek calculus~\citep{Lambek58}.
Fitch-style systems\todo{cite} probably also are precluded similarly.

>>>>

This chapter proceeds as follows.
In \cref{sec:semiring-motivation}, I give an intuitive introduction to semiring
annotations on variables, based on replicating features of DILL (introduced in
\cref{sec:dill}).
Then, \cref{sec:lr} formalises the ideas of \cref{sec:semiring-motivation} into
a calculus $\name$.
This calculus has
appeared in my previous work~\citep{WA21}, and can be seen as a
simply typed version of Atkey's dependently typed calculus QTT~\citep{Atkey18}.
%The idea of \name{} is to augment the simply typed $\lambda$-calculus with
%annotations on free variables, and give enough types to manipulate these
%annotations.
Given this new calculus $\name$, the first goal is to apply the techniques of
\cref{sec:simple} to it, yielding a simultaneous substitution operation.
To do this, I use \cref{sec:lnd} to introduce notation that allows us to restate
the typing rules of \name{} to not mention contexts explicitly, as was the style
in \cref{sec:simple}.
This new notation --- the \emph{bunched connectives} --- is versatile at
defining simply typed usage-aware syntaxes, and I give further non-$\name$
examples in \cref{sec:variant}.
Finally, I justify connections to linear logic and modal logic in
\cref{sec:translation}, where I translate $\name$ terms to and from
DILL~\citep{Barber1996} and the modal calculus of \citet{judgmental}.

\section{Motivation for semiring annotations}\label{sec:semiring-motivation}

The question of defining calculi which do not semantically admit weakening and
contraction but also do not rely on variables going out of scope is directly
addressed by \citet{McBride16}.
The first technique suggested is to, instead of
removing variables from the context of certain subterms, add an annotation to
free variables saying whether or not they are to be used.
I use an annotation $\gr0$ on variables that are not to be used, and an
annotation $\gr1$ on variables that are to be used.
This convention lets us transcribe the usual $\otimes$-introduction rule
(below left) as a rule with usage annotations (below right).
In the notation on the right, I let $\Gamma = \grP\gamma$ and
$\Delta = \grQ\delta$, where $\Gamma$ is a whole context comprising a
\emph{usage context} $\grP$ and a \emph{typing context} $\gamma$.
A usage context is a list of usage annotations, so
$\grP = \gr{r_1}, \ldots, \gr{r_m}$ and a typing context is a list of types, so
$\gamma = A_1, \ldots, A_m$.
When combined, the usage context and the typing context will be of the same
length.
Explicit contexts will usually be written with usage annotations and types
interspersed, as $\gr{r_1}A_1, \ldots, \gr{r_m}A_m$.
I use $\gr r\gamma$ to abbreviate $\gr rx_1, \ldots, \gr rx_m$.

@@

@@

An important point to note is that $\name$ cannot be instantiated to become a
bunched logic, in the sense of \citet{oHP99}.
This is despite the use of the bunched connectives in the \emph{definition} of
$\name$, and despite the formally tree-shaped contexts.
However, the variable rule of $\name$ essentially treats the variables in the
context independently, except for the individual checking of usage annotations
on each variable.
Therefore, we cannot talk about more interesting spatial relationships between
variables, as required by bunched logic.
This property also precludes us from capturing calculi without exchange
(non-commutative logics), such as Lambek calculus~\citep{Lambek58}.
Fitch-style systems~\citep{Borghuis-thesis} probably also are precluded
similarly.

### 2023-07-16 tex/simple.tex
[ writing ] address all remaining todos

In this chapter, I review and justify the family of approaches usually used to
represent simple type systems inside dependently typed proof assistants.
These approaches were first presented by \citet{AR99}, who showed a way of
representing well \emph{scoped} terms in a language with polymorphic recursion,
and extended the representation to well \emph{typed} terms in a language with
dependent types.
The representation of terms relies on indexing on both a context --- giving the
types of all the free variables --- and a type for the term itself.
A basic operation on terms is \emph{simultaneous substitution}, which replaces
each variable in the context by a term in another context.
This simultaneous substitution operation fits the form of composition in a
multicategory.
Specifically, in the case of a simple intuitionistic type system, this is the
composition of a Cartesian multicategory, which allows all of the terms being
substituted in to share a context.
\todo{Maybe delete multicategory remark.}

This chapter almost entirely reviews prior work.
In \cref{sec:terms}, I use a presentation of the dependently typed encoding of
the simply typed $\lambda$-calculus from \citet{AR99} to set notational
conventions.
\Cref{sec:kits} presents the unpublished work of \citet{McBride05},
successively deriving simultaneous renaming and simultaneous substitution for
the terms defined in \cref{sec:terms}.
The rest of the chapter generalises the shared core of renaming and substitution
in two dimensions: in \cref{sec:gen-sem} following \citet{ACMM17} to cover
semantic traversals, and in \cref{sec:gen-syn} following \citet{AACMM21} to
cover a whole range of simply typed syntaxes with binding, rather than just
a specific syntax.
Finally, I review some related work in \cref{sec:mech-related}.

\section{Term representation}\label{sec:terms}
\input{terms.tex}
\section{Renaming and substitution}\label{sec:kits}
\input{kits.tex}
\section{Generic semantics}\label{sec:gen-sem}
\input{gen-sem.tex}
\section{Generic syntax}\label{sec:gen-syn}
\input{gen-syn.tex}
%\section{Natural deduction vs sequent calculus}
%\input{ndsc.tex}
\section{Related work}\label{sec:mech-related}
\input{mech-related.tex}

>>>>

In this chapter, I review and justify the family of approaches usually used to
represent simple type systems inside dependently typed proof assistants.
These approaches were first presented by \citet{AR99}, who showed a way of
representing well \emph{scoped} terms in a language with polymorphic recursion,
and extended the representation to well \emph{typed} terms in a language with
dependent types.
The representation of terms relies on indexing on both a context --- giving the
types of all the free variables --- and a type for the term itself.
A basic operation on terms is \emph{simultaneous substitution}, which replaces
each variable in the context by a term in another context.
%This simultaneous substitution operation fits the form of composition in a
%multicategory, giving us a connection to categorical logic~\citep{}.
%Specifically, in the case of a simple intuitionistic type system, this is the
%composition of a Cartesian multicategory, which allows all of the terms being
%substituted in to share a context.

This chapter almost entirely reviews prior work.
First, in \cref{sec:agda-primer}, I give an introduction to Agda, the proof
assistant I work in throughout this thesis.
Then, in \cref{sec:terms}, I use a presentation of the dependently typed
encoding of the simply typed $\lambda$-calculus from \citet{AR99} to set
notational conventions.
\Cref{sec:kits} presents the unpublished work of \citet{McBride05},
successively deriving simultaneous renaming and simultaneous substitution for
the terms defined in \cref{sec:terms}.
The rest of the chapter generalises the shared core of renaming and substitution
in two dimensions: in \cref{sec:gen-sem} following \citet{ACMM17} to cover
semantic traversals, and in \cref{sec:gen-syn} following \citet{AACMM21} to
cover a whole range of simply typed syntaxes with binding, rather than just
a specific syntax.
Finally, I review some related work in \cref{sec:mech-related}.

\input{agda-primer.tex}
\section{Term representation}\label{sec:terms}
\input{terms.tex}
\section{Renaming and substitution}\label{sec:kits}
\input{kits.tex}
\section{Generic semantics}\label{sec:gen-sem}
\input{gen-sem.tex}
\section{Generic syntax}\label{sec:gen-syn}
\input{gen-syn.tex}
%\section{Natural deduction vs sequent calculus}
%\input{ndsc.tex}
\section{Related work}\label{sec:mech-related}
\input{mech-related.tex}
